Online {\IterComp}

### 2017-11-17 motivation.tex
overall changes

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose. To show this, we
    performed iterative compilation on \texttt{tiff2bw}, evaluating each optimization sequence on a different input and using either
    runtime or IPC as our efficiency metric. We selected a set of ten inputs from the KDataSets. When
    iterative compilation is driven by runtime, the selected optimization slows down the benchmark by 3\% compared to \texttt{-O3}. When
    driven by IPC, it is even worse with an average slowdown of 49\%. Its very high IPC of 3.25 does not indicate high performance. The
    actual best optimization over the whole input set improves performance by 7\% while only having an average IPC of 2.31.

\begin{table}[h]
\centering
\label{tab:motivation_metrics}
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{1}{|c|}{Metric} & \multicolumn{1}{c|}{Speedup/Slowdown} & \multicolumn{1}{c|}{IPC} \\ \hline
Oracle (Speedup)             & 7\% speedup                                 & 2.31                      \\ \hline
IPC                          & 49\% slowdown                                & 3.25                      \\ \hline
1/Runtime                    & 3\% slowdown                                  & \multicolumn{1}{c|}{-}   \\ \hline
\end{tabular}
\caption{}
\end{table}

>>>>

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose. To show this, we
    performed iterative compilation on \texttt{tiff2bw}, evaluating each optimization sequence on a different input and using either
    runtime or IPC as our efficiency metric. We selected a set of ten inputs from the KDataSets. When
    iterative compilation is driven by runtime, the selected optimization slows down the benchmark by 3\% compared to \texttt{-O3}. When
    driven by IPC, it is even worse with an average slowdown of 49\%. Its very high IPC of 3.25 does not indicate high performance. The
    actual best optimization over the whole input set improves performance by 7\% while only having an average IPC of 2.31.
   Higher IPC does not necessarily translate to better performance,
 different optimizations affect IPC in different ways.
   Similarly, a shorter runtime may be due to inputs with fewer work for the program to process.

\begin{table}[h]
\centering
\label{tab:motivation_metrics}
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{1}{|c|}{Metric} & \multicolumn{1}{c|}{Speedup/Slowdown} & \multicolumn{1}{c|}{IPC} \\ \hline
Oracle (Speedup)             & 7\% speedup                                 & 2.31                      \\ \hline
IPC                          & 49\% slowdown                                & 3.25                      \\ \hline
1/Runtime                    & 3\% slowdown                                  & \multicolumn{1}{c|}{-}   \\ \hline
\end{tabular}
\caption{Comparison of iterative compilation guided by different metrics.
The oracle is guided by the average speedup over \texttt{-O3} on all inputs.
For IPC and runtime, each optimization was evaluated on a different input.
}
\vspace{-2ex}
\end{table}

### 2017-11-17 results.tex
overall changes

\subsubsection{Evaluating Costs During the Online Search}

\begin{figure*}[t]
  \centering
  \subfigure[Opt-Profiling]{\includegraphics[width=0.3\textwidth]{figs/Opt-Profiling-deployment.pdf}}
  \subfigure[WC-Relax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WC-Relax-5-deployment.pdf}}
  \subfigure[WP-Relax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WP-Relax-5-deployment.pdf}}
  \caption{.}
  \label{fig:deployment_cost}
\end{figure*}

We also evaluate the actual cost perceived by the user when applying online iterative compilation.
To this end we used the architecture described in Section~\ref{sec:oic-infra} with the search based on genetic algorithm.
Our results provide evidence that the proposed approach is profitable in real-world online scenarios.
Figure~\ref{fig:deployment_cost} shows, for each benchmark and configuration, the profitability of the final optimization sequence
selected by the search (as in Figure~\ref{fig:speedups}) but also the average profitability/cost experienced by the user throughout the search process (labeled as [Live]),
which takes into account the overhead of the work profiling.

With online iterative optimization, although the 
user may experience slowdown during the search process by executing sub-optimal versions of the program,
a Search Driver based on genetic algorithm is able to focus the search on promising optimization sequences, reducing the risk of
executing sub-optimal versions of the program as the search evolves through generations.
After the first generation, the optimization sequences produced tend to overcome this issue, with their benefits surpassing the profiling overheads.

For the \WCRelax-5\% configuration, an average speedup of XX was observed during the search, in the worst case with an slowdown of XX.
For the \WPRelax-5\% configuration, an average speedup of XX was observed during the search, in the worst case with an slowdown of XX.

>>>>

\subsubsection{Cost Evaluation During the Online Search}

\begin{figure*}[t]
  \centering
  \subfigure[\OptProf]{\includegraphics[width=0.3\textwidth]{figs/Opt-Profiling-deployment.pdf}}
  \subfigure[\WCRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WC-Relax-5-deployment.pdf}}
  \subfigure[\WPRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WP-Relax-5-deployment.pdf}}
  \caption{Evaluation of the profitability/cost experienced by the user throughout the search process (labeled as [Live]) and of the best optimization sequence selected by the search. These results provide evidence that the proposed approach is profitable in real-world online scenarios.}
  \label{fig:deployment_cost}
\end{figure*}

We also evaluate the actual cost perceived by the user when applying online iterative compilation.
To this end we used the architecture described in Section~\ref{sec:oic-infra} with the search based on genetic algorithm~\cite{knijnenburg02,cooper99,kulkarni04}.
Our results provide evidence that the proposed approach is profitable in real-world online scenarios.
Figure~\ref{fig:deployment_cost} shows, for each benchmark and configuration, the profitability of the final optimization sequence
selected by the search (as in Figure~\ref{fig:speedups}) but also the average profitability/cost experienced by the user throughout the search process (labeled as [Live]),
which takes into account the overhead of the work profiling.

With online iterative optimization, although the 
user may experience slowdown during the search process by executing sub-optimal versions of the program,
a Search Driver based on genetic algorithm is able to focus the search on promising optimization sequences, reducing the risk of
executing sub-optimal versions of the program as the search evolves through generations.
After the first generation, the optimization sequences produced tend to overcome this issue, with their benefits surpassing the profiling overheads.
We evaluate the deployment cost for the three configurations that use work profiling, namely \OptProf, \WCRelax-5\%, and \WPRelax-5\%.
In all three configurations, the online search has been shown profitable on average.
The wider search space is also able to improve the performance gains over \texttt{-O3}.
